MLH1 Results Outline

### 2020-02-20 doc/Results.rmd
small formating edits

#Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
#Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

@@

The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific=F )`, and `r Q1_tot.sc_t.test.LEW$p.value`
for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _ respectively).

Some of the p values for PWD and MSM are much closer (less significant), ( p = `r Q1_tot.sc_t.test.PWD$p.value`, `r Q1_tot.sc_t.test.KAZ$p.value`, and `r Q1_tot.sc_t.test.MSM$p.value` for the _M. m. musculus^PWD^ _, _M. m. musculus^KAZ^ _, and _M. m. molossinus^MSM^_ respectively)

@@

```{r Q1.lmer_totalSC, echo=FALSE}

@@

Q1.lmer.totSC_results

#G
```
there are more significant effects for strain and subsp for the mixed model of mouse.av total sc

$$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$

@@

#G * malev.sig and G*LEW

@@

@@

When skive is plotted alone it does have overlapping ranges.

@@

(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
for WSB, G and LEW respectively.

(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value` for PWD, KAZ, SKIVE

@@

$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1_short_biv_lmer, echo=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)
Q1lmer.MM_SC_short_results

Sex is the most significant factor for the mixed model of mean short bivalents.

$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.SC.Length_M2, echo=FALSE}
Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)

$$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$

@@

@@

we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference. We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ). (all chromosomes in house mouse are acrocentric)

@@

```{r CO.positions.figs, echo=FALSE, include=FALSE}
#main figure of density plots,

>>>>

@@

The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).

Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)

@@

```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE}

@@

#Q1.lmer.totSC_results

#G
```

there are more significant effects for strain and subsp for the mixed model of mouse.av total sc

the 3 fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`)

These effects were followed up the glms

**M2.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$

**M3.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$

@@

#G * male v.sig and male*LEW

@@

The results from those models was that there were 
...lots of significant effects.  (G male, )

The summary for the total SC area comparison is...

<!-- transition to short-->

The short bivalent dataset was used -- because there are no confounding factors of the XX -- so makes hapier comparisons across male and females

@@

@@

(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
for WSB, G and LEW respectively.

(t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for PWD, KAZ, SKIVE

@@

**M2.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE}

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)

**M3.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)

#Q1lmer.MM_SC_short_results

Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). interaction effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`)

**M3.** $$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$

@@

In summary .... females (oocytes) have different chromatin compaction patterns compared to males.

@@

We focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference. We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ). (all chromosomes in house mouse are acrocentric)

@@

```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
#main figure of density plots,
